\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-11-29 doc/paper/paper.tex
Some minor rewording, thanks Luis for the feedback.

Fig.~\ref{lst:vigor} shows the annotated code, ready for the verification with
Vigor. It is the result of applying the first two steps shown below.
\code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
would like to verify --- that we do not forward packets with port 9.

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
  if (!r) return 1;
  while(LOOP(1))^\label{line:loop-hd}^
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)^\label{line:if-cond}^
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);^\label{line:pop-front}^
      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.
    \code{LOOP} (l.\ref{line:loop-hd}) labels the infinite loop head and we
    explicitly use a ring (l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
    line \ref{line:vigcheck} adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

@@

{\bf Step 2: Use library components. } A developer replaces every array with a
component from Vigor library that performs the same function. For our example,
the library provides a packet ring with a simple interface: \code{bool
  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
  ring_full(ring*)} checks whether the ring has run out of space, \code{void
  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
and places it in \code{dst}.

{\bf Step 3: Provide invariants.} Besides usual parameters, like capacity or
element size, most Vigor components also require simple logical invariants of
their elements to enable the reasoning about the application code. For our
example, the essential property of the ring is, that it contains only packets
with a port different from 9. A developer communicates that as shown in
Fig.~\ref{lst:customizations}.

@@

Vigor transforms infinite loops using loop
invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
automatic havoc-ing~\cite{barnett2005boogie}. Havoc-ing simply replaces all the
variables that may change in the loop by unconstrained symbols at the beginning
of the iteration. In our example that would be the variables \code{begin} and
\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
of the \code{packets} pointer because it never changes. The step 1 enables
Vigor to reduce each problematic loop to a single generalized iteration that
allows reasoning in one pass.

@@

In Vigor, we circumvent the problem of symbolic indexing by encapsulating
program state into well-defined data structures and putting them into a
``library of components'' (e.g. the ring in the step 2). This works well for NF
development, because the NF code uses a relatively small set of standard
structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
We provide to the SEE symbolic models of these abstracted components that
``hide'' symbolic indexes and pointers from the SEE.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));^\label{line:assume}^
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
    a packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} that is inserted automatically at
the beginning of a loop iteration. It continues then with calls to
\code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
return \code{false}. They both return false in this particular trace, in other
traces they returned other values, but to define the context for the last
function call (l.\ref{line:tip-call}) we must repeat the history as closely as
possible. Vigor handles the result of the last call in the sequence very
differently. Instead of just telling the formal verification engine what
happened, it inserts an \code{assert}(l.\ref{line:assert}) reflecting the
symbolic model output to check that it is consistent with the postcondition of
the last function (Fig.~\ref{lst:contract}).

@@

/*@ assert(arg2.port != 9); @*/^\label{line:assert}^
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space. Unexplained
    lines(\ref{line:unexplained1},\ref{line:unexplained2},\ref{line:unexplained3},\ref{line:unexplained4})
    are the automatically inserted helper steps to guide the proof checker.}
  \label{lst:call-sequence}
\end{figure}

Formal verification engine checks this sequence using the contracts like in the
Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g. call for
\code{ring_pop} on a possibly empty ring. It then checks that the model output
covers all the possible outputs specified in the contract for the tip function.
This happens in the last line on the listing. The assert contains the
path-constraint generated by the symbolic model. In this case the contract
(Fig.\ref{lst:contract}) guarantees that the popped packet satisfies
\code{user_property} that entails \code{arg2.port != 9}, hence the trace is
legitimate. (see~\S\ref{sec:model-validation-criteria}).

@@

\subsection{Model inconsystency symptoms}

>>>>

Fig.~\ref{lst:vigor} shows the annotated code, ready for the verification with
Vigor. It is the result of applying the first two steps shown below.
\code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
would like to verify: ``we do not forward packets destined to port 9''.

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
  if (!r) return 1;
  while(LOOP(1))^\label{line:loop-hd}^
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)^\label{line:if-cond}^
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);^\label{line:pop-front}^
      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Annotated code of Drop buffer.
    \code{LOOP} (l.\ref{line:loop-hd}) labels the infinite loop head and we
    explicitly use a ring (l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
    line \ref{line:vigcheck} adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

@@

{\bf Step 2: Use library components. } A developer replaces state chunks (e.g.
most of the arrays) with components from the Vigor library that perform the
same function. For our example, the library provides a packet ring with a simple
interface: \code{bool ring_empty(ring*)} checks whether there is a packet in the
ring, \code{bool ring_full(ring*)} checks whether the ring has run out of space,
\code{void ring_push_back(ring*, packet* src)} adds packet \code{src} to the
ring, and \code{void ring_pop_front(ring*, packet* dst)} removes a packet from
the ring and places it in \code{dst}.

{\bf Step 3: Provide invariants.} Besides usual parameters, like capacity or
element size, most Vigor components also require simple logical invariants about
their elements to enable the reasoning about the application code. For our
example, the essential property of the ring is, that it contains only packets
with a port different from 9. A developer communicates that as shown in
Fig.~\ref{lst:customizations}.

@@

Vigor transforms infinite loops using loop
invariants~\cite[\S~2.1]{cormen2009introduction} that allow to model all the
iterations of a loop in a single one. To support this kind of generalization, we
added automatic havoc-ing~\cite{barnett2005boogie} to KLEE. Havoc-ing simply
replaces all the variables that may change in the loop by unconstrained symbols
at the beginning of the iteration. In our example that would be the variables
\code{begin} and \code{queue_len} and the contents of \code{packets[0..CAP]},
but not the value of the \code{packets} pointer because it never changes. The
step 1 enables Vigor to reduce each problematic loop to a single generalized
iteration that allows reasoning in one pass.

@@

In Vigor, we circumvent the problem of symbolic indexing by encapsulating
program state into well-defined data structures and putting them into a
``library of components'' (e.g. the ring in step 2). This works well for NF
development, because the NF code uses a relatively small set of standard
structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
We provide to the SEE symbolic models of these abstracted components that
``hide'' symbolic indexes and pointers from the SEE.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));^\label{line:assume}^
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically a
    packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} and continues then with calls to
\code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
return \code{false}. They both return false in this particular trace, in other
traces they returned other values, but to define the context for the last
function call (l.\ref{line:tip-call}) we must repeat the history as closely as
possible. Vigor handles the result of the last call in the sequence very
differently. Instead of just telling the formal verification engine what
happened, it inserts an \code{assert}(l.\ref{line:assert}) reflecting the
symbolic model output to check that it is consistent with the postcondition of
the last function (Fig.~\ref{lst:contract}).

@@

/*@ assert(arg2.port != 9); @*/^\label{line:assert}^
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space. The \code{loop_invariant_produce} call is
    inserted automatically at the beginning of a loop iteration. Lines
    \ref{line:unexplained1},\ref{line:unexplained2},\ref{line:unexplained3},\ref{line:unexplained4}
    account for the automatically inserted helper steps to guide the proof
    checker.}
  \label{lst:call-sequence}
\end{figure}

The formal verification engine checks this sequence using contracts like the one
shown in the Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g.
call for \code{ring_pop} on a possibly empty ring. It then checks that the model
output covers all the possible outputs specified in the contract for the last
function. This happens in the last line on the listing. The assert contains the
path-constraint generated by the symbolic model. In this case the contract
(Fig.\ref{lst:contract}) guarantees that the popped packet satisfies
\code{user_property} that entails \code{arg2.port != 9}, hence the trace is
legitimate. (see~\S\ref{sec:model-validation-criteria}).

@@

\subsection{Detecting model inconsistencies}
